Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Berlin version \YellowPaperVersionNumber}}

### 2022-01-16 Paper.tex
EIP-2930: Optional access lists

EIP-2930 (type 1) transactions also have:
\begin{description}
\item[accessList]\linkdest{tx_access_list}{} TODO; formally $T_{\mathbf{a}}$.
\item[chainId]\linkdest{tx_chain_id}{} Chain ID; formally $T_{\mathrm{c}}$. Must be equal to the network chain ID \hyperlink{chain_id}{$\beta$}.
\item[yParity]\linkdest{tx_y_parity}{} Signature Y parity; formally $T_{\mathrm{y}}$.
\end{description}

Legacy transactions do not have an \textbf{accessList}, while \textbf{chainId} and \textbf{yParity} for legacy transactions are combined into a single value:
\begin{description}
\item[w]\linkdest{T__w}{} A scalar value encoding Y parity and possibly chain ID; formally $T_{\mathrm{w}}$.
$T_{\mathrm{w}} = 27 + T_{\mathrm{y}}$ or $T_{\mathrm{w}} = 2\beta + 35 + T_{\mathrm{y}}$ (see EIP-155 by \cite{EIP-155}).
\end{description}

@@

\begin{equation}
\linkdest{L_transaction} L_{\mathrm{T}}(T) \equiv \begin{cases}
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 0 \\
(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{a}}, T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 1 \\
\end{cases}
\end{equation}
where
\begin{equation}
\mathbf{p} \equiv \begin{cases}
T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = \varnothing \\
T_{\mathbf{d}} & \text{otherwise}
\end{cases}
\end{equation}

Here, we assume all components are interpreted by the RLP as integer values, with the exception of the access list $T_{\mathbf{a}}$ and the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
\begin{equation}
\begin{array}[t]{lclclc}
T_{\mathrm{x}} \in \{0, 1\} & \wedge & T_{\mathrm{c}} = \beta & \wedge & T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge \\
T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge \\
T_{\mathrm{w}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge \\
T_{\mathrm{y}} \in \mathbb{N}_{1} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
\end{array}
\end{equation}
where
\begin{equation}
\mathbb{N}_{\mathrm{n}} = \{ P: P \in \mathbb{N} \wedge P < 2^n \}
\end{equation}

@@

\subsection{Substate} \label{ch:substate}
Throughout transaction execution, we accrue certain information that is acted upon immediately following the transaction. We call this the \textit{accrued transaction substate}, or \textit{accrued substate} for short, and represent it as $A$, which is a tuple:
\begin{equation}
A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}}, A_{\mathbf{a}}, A_{\mathbf{k}})
\end{equation}

\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
\hypertarget{refund_balance_defn_words_A__r}{}$A_{\mathrm{r}}$ is the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value. Though not immediately refunded, it is allowed to partially offset the total execution costs.
Finally, EIP-2929 by \cite{EIP-2929} introduced \hypertarget{accessed_addresses_defn_words_A__a}{}$A_{\mathbf{a}}$, the set of accessed account addresses, and \hypertarget{accessed_storage_keys_defn_words_A__k}{}$A_{\mathbf{k}}$, the set of accessed storage keys.

@@

\subsection{Execution}
\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
\begin{align}
g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
\nonumber {} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
\nonumber {} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}}
\end{align}
where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call. $G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code. $G$ is fully defined in Appendix \ref{app:fees}.

@@

Evaluating $\boldsymbol{\sigma}_{\mathrm{P}}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
\begin{equation}
(\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
\hyperlink{lambda}{\Lambda}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), g, &\\ \quad\; T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
\hyperlink{theta}{\Theta}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), T_{\mathrm{t}}, &\\ \quad\; T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
\end{cases}
\end{equation}
where
\begin{equation}
A^* \equiv A^0 \quad \text{except} \quad A^*_{\mathbf{a}} \equiv A^0_{\mathbf{a}} \cup \{S(T)\}
\end{equation}
and $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
\begin{equation}
g \equiv T_{\mathrm{g}} - g_0
\end{equation}

@@

\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak-256 hash of the transaction. Three different flavours of signing schemes are available:
\begin{eqnarray}
L_{\mathrm{X}}(T) & \equiv & \begin{cases}
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{27, 28\} \\
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \beta, (), ()) & \text{if} \; T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{2\beta + 35, 2\beta + 36\} \\
(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{a}}) & \text{if} \; T_{\mathrm{x}} = 1
\end{cases} \\
\nonumber \text{where} \\
\nonumber \mathbf{p} & \equiv & \begin{cases}
T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = \varnothing \\
T_{\mathbf{d}} & \text{otherwise}
\end{cases} \\
h(T) & \equiv & \begin{cases}
\mathtt{KEC}( \mathtt{RLP}(L_{\mathrm{X}}(T)) ) & \text{if} \; T_{\mathrm{x}} = 0 \\
\mathtt{KEC}( T_{\mathrm{x}} \cdot \mathtt{RLP}(L_{\mathrm{X}}(T)) ) & \text{otherwise}
\end{cases}
\end{eqnarray}

@@

\nopagebreak
\begin{tabu}{l r l}
\toprule
Name & Value & Description \\
\midrule
$G_{\mathrm{zero}}$ & 0 & Nothing paid for operations of the set {\small $W_{\mathrm{zero}}$}. \\
$G_{\mathrm{jumpdest}}$ & 1 & Amount of gas to pay for a {\small JUMPDEST} operation. \\
$G_{\mathrm{base}}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{base}}$}. \\
$G_{\mathrm{verylow}}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{verylow}}$}. \\
$G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
$G_{\mathrm{mid}}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{mid}}$}. \\
$G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
$G_{\mathrm{warmaccess}}$ & 100 & Cost of a warm account or storage access. \\
$G_{\mathrm{accesslistaddress}}$ & 2400 & Cost of warming up an account with the access list. \\
$G_{\mathrm{accessliststorage}}$ & 1900 & Cost of warming up a storage with the access list. \\
$G_{\mathrm{coldaccountaccess}}$ & 2600 & Cost of a cold account access. \\
$G_{\mathrm{coldsload}}$ & 2100 & Cost of a cold storage access. \\
$G_{\mathrm{sset}}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
$G_{\mathrm{sreset}}$ & 2900 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
&&is set to zero. \\
$R_{\mathrm{sclear}}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
&&non-zero.\\
\linkdest{R__selfdestruct}{}$R_{\mathrm{selfdestruct}}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
\linkdest{G__selfdestruct}{}$G_{\mathrm{selfdestruct}}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
$G_{\mathrm{create}}$ & 32000 & Paid for a {\small CREATE} operation. \\
\linkdest{G__codedeposit}{}$G_{\mathrm{codedeposit}}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
$G_{\mathrm{callvalue}}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
\linkdest{G__callstipend}{}$G_{\mathrm{callstipend}}$ & 2300 & A stipend for the called contract subtracted from $G_{\mathrm{callvalue}}$ for a non-zero value transfer. \\
\linkdest{G__newaccount}{}$G_{\mathrm{newaccount}}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
$G_{\mathrm{exp}}$ & 10 & Partial payment for an {\small EXP} operation. \\
$G_{\mathrm{expbyte}}$ & 50 & Partial payment when multiplied by the number of bytes in the exponent for the {\small EXP} operation. \\
$G_{\mathrm{memory}}$ & 3 & Paid for every additional word when expanding memory. \\
\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
\linkdest{G__txdatazero}{}$G_{\mathrm{txdatazero}}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
\linkdest{G__txdatanonzero}{}$G_{\mathrm{txdatanonzero}}$ & 16 & Paid for every non-zero byte of data or code for a transaction. \\
\linkdest{G__transaction}{}$G_{\mathrm{transaction}}$ & 21000 & Paid for every transaction. \\
$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
$G_{\mathrm{keccak256}}$ & 30 & Paid for each {\small KECCAK256} operation. \\
$G_{\mathrm{keccak256word}}$ & 6 & Paid for each word (rounded up) for input data to a {\small KECCAK256} operation. \\
$G_{\mathrm{copy}}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
$G_{\mathrm{blockhash}}$ & 20 & Payment for each {\small BLOCKHASH} operation. \\

@@

&&&& $A'_{\mathbf{k}} \equiv A_{\mathbf{k}} \cup \{(I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0])\}$ \\

@@

G_{\mathrm{warmaccess}} & \text{if} \quad (I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0]) \in A_{\mathbf{k}} \\

@@

0                      & \text{if} \quad (I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0]) \in A_{\mathbf{k}} \\

>>>>

EIP-2930 (type 1) transactions also have:
\begin{description}
\item[accessList]\linkdest{tx_access_list}{} List of access entries to warm up; formally $T_{\mathbf{A}}$.
Each access list entry $A$ is a tuple of an account address and a list of storage keys: $A \equiv (A_{\mathrm{a}}, A_{\mathbf{s}})$. 
\item[chainId]\linkdest{tx_chain_id}{} Chain ID; formally $T_{\mathrm{c}}$. Must be equal to the network chain ID \hyperlink{chain_id}{$\beta$}.
\item[yParity]\linkdest{tx_y_parity}{} Signature Y parity; formally $T_{\mathrm{y}}$.
\end{description}

Legacy transactions do not have an \textbf{accessList} ($T_{\mathbf{A}}=()$), while \textbf{chainId} and \textbf{yParity} for legacy transactions are combined into a single value:
\begin{description}
\item[w]\linkdest{T__w}{} A scalar value encoding Y parity and possibly chain ID; formally $T_{\mathrm{w}}$.
$T_{\mathrm{w}} = 27 + T_{\mathrm{y}}$ or $T_{\mathrm{w}} = 2\beta + 35 + T_{\mathrm{y}}$ (see EIP-155 by \cite{EIP-155}).
\end{description}

@@

\begin{equation}
\linkdest{L_transaction} L_{\mathrm{T}}(T) \equiv \begin{cases}
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 0 \\
(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}, T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 1 \\
\end{cases}
\end{equation}
where
\begin{equation}
\mathbf{p} \equiv \begin{cases}
T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = \varnothing \\
T_{\mathbf{d}} & \text{otherwise}
\end{cases}
\end{equation}

Here, we assume all components are interpreted by the RLP as integer values, with the exception of the access list $T_{\mathbf{A}}$ and the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
\begin{equation}
\begin{array}[t]{lclclc}
T_{\mathrm{x}} \in \{0, 1\} & \wedge & T_{\mathrm{c}} = \beta & \wedge & T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge \\
T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge \\
T_{\mathrm{w}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge \\
T_{\mathrm{y}} \in \mathbb{N}_{1} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
\end{array}
\end{equation}
where
\begin{equation}
\mathbb{N}_{\mathrm{n}} = \{ P: P \in \mathbb{N} \wedge P < 2^n \}
\end{equation}

@@

\subsection{Substate} \label{ch:substate}
Throughout transaction execution, we accrue certain information that is acted upon immediately following the transaction. We call this the \textit{accrued transaction substate}, or \textit{accrued substate} for short, and represent it as $A$, which is a tuple:
\begin{equation}
A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}}, A_{\mathbf{a}}, A_{\mathbf{K}})
\end{equation}

\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
\hypertarget{refund_balance_defn_words_A__r}{}$A_{\mathrm{r}}$ is the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value. Though not immediately refunded, it is allowed to partially offset the total execution costs.
Finally, EIP-2929 by \cite{EIP-2929} introduced \hypertarget{accessed_addresses_defn_words_A__a}{}$A_{\mathbf{a}}$, the set of accessed account addresses, and \hypertarget{accessed_storage_keys_defn_words_A__k}{}$A_{\mathbf{K}}$, the set of accessed storage keys
(more accurately, each element of $A_{\mathbf{K}}$ is a tuple of a 20-byte account address and a 32-byte storage slot).

@@

\subsection{Execution}
\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
\begin{align}
g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
\nonumber {} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
\nonumber {} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}} \\
\nonumber {} & + \sum_{j = 0}^{\lVert T_{\mathbf{A}} \rVert - 1} \big( G_{\mathrm{accesslistaddress}} + \lVert T_{\mathbf{A}}[j]_{\mathbf{s}} \rVert G_{\mathrm{accessliststorage}} \big)
\end{align}
where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
$\hyperlink{G__accesslistaddress}{G_{\mathrm{accesslistaddress}}}$ and $\hyperlink{G__accessliststorage}{G_{\mathrm{accessliststorage}}}$ are the costs of warming up account and storage access, respectively.
$G$ is fully defined in Appendix \ref{app:fees}.

@@

Evaluating $\boldsymbol{\sigma}_{\mathrm{P}}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
\begin{equation}
(\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
\hyperlink{lambda}{\Lambda}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), g, &\\ \quad\; T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
\hyperlink{theta}{\Theta}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), T_{\mathrm{t}}, &\\ \quad\; T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
\end{cases}
\end{equation}
where
\begin{align}
A^* & \equiv A^0 \quad \text{except} \\
A^*_{\mathbf{a}} & \equiv A^0_{\mathbf{a}} \cup \{S(T)\} \cup \{ T_{\mathbf{A}}[i]_{\mathrm{a}}, \; i < \lVert T_{\mathbf{A}} \rVert \} \\
A^*_{\mathbf{K}} & \equiv A^0_{\mathbf{K}} \bigcup_{i = 0}^{\lVert T_{\mathbf{A}} \rVert - 1} \{ (T_{\mathbf{A}}[i]_{\mathrm{a}}, T_{\mathbf{A}}[i]_{\mathbf{s}}[j]), \; j < \lVert T_{\mathbf{A}}[i]_{\mathbf{s}} \rVert \}
\end{align}
and $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
\begin{equation}
g \equiv T_{\mathrm{g}} - g_0
\end{equation}

@@

\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak-256 hash of the transaction. Three different flavours of signing schemes are available:
\begin{eqnarray}
L_{\mathrm{X}}(T) & \equiv & \begin{cases}
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{27, 28\} \\
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \beta, (), ()) & \text{if} \; T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{2\beta + 35, 2\beta + 36\} \\
(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}) & \text{if} \; T_{\mathrm{x}} = 1
\end{cases} \\
\nonumber \text{where} \\
\nonumber \mathbf{p} & \equiv & \begin{cases}
T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = \varnothing \\
T_{\mathbf{d}} & \text{otherwise}
\end{cases} \\
h(T) & \equiv & \begin{cases}
\mathtt{KEC}( \mathtt{RLP}(L_{\mathrm{X}}(T)) ) & \text{if} \; T_{\mathrm{x}} = 0 \\
\mathtt{KEC}( T_{\mathrm{x}} \cdot \mathtt{RLP}(L_{\mathrm{X}}(T)) ) & \text{otherwise}
\end{cases}
\end{eqnarray}

@@

\nopagebreak
\begin{tabu}{l r l}
Name & Value & Description \\
$G_{\mathrm{zero}}$ & 0 & Nothing paid for operations of the set {\small $W_{\mathrm{zero}}$}. \\
$G_{\mathrm{jumpdest}}$ & 1 & Amount of gas to pay for a {\small JUMPDEST} operation. \\
$G_{\mathrm{base}}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{base}}$}. \\
$G_{\mathrm{verylow}}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{verylow}}$}. \\
$G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
$G_{\mathrm{mid}}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{mid}}$}. \\
$G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
$G_{\mathrm{warmaccess}}$ & 100 & Cost of a warm account or storage access. \\
\linkdest{G__accesslistaddress}{}$G_{\mathrm{accesslistaddress}}$ & 2400 & Cost of warming up an account with the access list. \\
\linkdest{G__accessliststorage}{}$G_{\mathrm{accessliststorage}}$ & 1900 & Cost of warming up a storage with the access list. \\
$G_{\mathrm{coldaccountaccess}}$ & 2600 & Cost of a cold account access. \\
$G_{\mathrm{coldsload}}$ & 2100 & Cost of a cold storage access. \\
$G_{\mathrm{sset}}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
$G_{\mathrm{sreset}}$ & 2900 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
&&is set to zero. \\
$R_{\mathrm{sclear}}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
&&non-zero.\\
\linkdest{R__selfdestruct}{}$R_{\mathrm{selfdestruct}}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
\linkdest{G__selfdestruct}{}$G_{\mathrm{selfdestruct}}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
$G_{\mathrm{create}}$ & 32000 & Paid for a {\small CREATE} operation. \\
\linkdest{G__codedeposit}{}$G_{\mathrm{codedeposit}}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
$G_{\mathrm{callvalue}}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
\linkdest{G__callstipend}{}$G_{\mathrm{callstipend}}$ & 2300 & A stipend for the called contract subtracted from $G_{\mathrm{callvalue}}$ for a non-zero value transfer. \\
\linkdest{G__newaccount}{}$G_{\mathrm{newaccount}}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
$G_{\mathrm{exp}}$ & 10 & Partial payment for an {\small EXP} operation. \\
$G_{\mathrm{expbyte}}$ & 50 & Partial payment when multiplied by the number of bytes in the exponent for the {\small EXP} operation. \\
$G_{\mathrm{memory}}$ & 3 & Paid for every additional word when expanding memory. \\
\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
\linkdest{G__txdatazero}{}$G_{\mathrm{txdatazero}}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
\linkdest{G__txdatanonzero}{}$G_{\mathrm{txdatanonzero}}$ & 16 & Paid for every non-zero byte of data or code for a transaction. \\
\linkdest{G__transaction}{}$G_{\mathrm{transaction}}$ & 21000 & Paid for every transaction. \\
$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
$G_{\mathrm{keccak256}}$ & 30 & Paid for each {\small KECCAK256} operation. \\
$G_{\mathrm{keccak256word}}$ & 6 & Paid for each word (rounded up) for input data to a {\small KECCAK256} operation. \\
$G_{\mathrm{copy}}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
$G_{\mathrm{blockhash}}$ & 20 & Payment for each {\small BLOCKHASH} operation. \\

@@

&&&& $A'_{\mathbf{K}} \equiv A_{\mathbf{K}} \cup \{(I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0])\}$ \\

@@

G_{\mathrm{warmaccess}} & \text{if} \quad (I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0]) \in A_{\mathbf{K}} \\

@@

0                      & \text{if} \quad (I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0]) \in A_{\mathbf{K}} \\
